\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species

### 2024-03-13 ms.Rnw
Update

<<Setup, include=FALSE,echo=FALSE,results=hide>>=
 library(plyr)
 library(reshape2)
 library(magrittr)
 library(foreach)
 library(readxl)
 library(doParallel)
 registerDoParallel(cores=36)
 library(pomp)
 library(panelPomp)
 library(ggplot2)

@@

 #---------------------------Host Daphnia----------------------------------------------------------------------
 #----------------load data---------------------------------------------------------------
load("/Users/ybb/Desktop/Daphnia-article/data/raw_data.RData")
 #-----------------------------Host with para----------------------------------------------------------------------------
 host_para <- Mesocosm_data[101:180, ]
 host_para <- subset(host_para, select = c(rep, day, dent.adult,dent.inf ))
 host_para <- host_para[80: 1, ]
 #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
 host_para$day = (host_para$day - 1) * 5 + 7
 host_para_data = list()
 trails = c("K","L","M","N","O","P","Q","S")
 for (i in 1: length(trails)){
   host_para_data[[i]]=subset(host_para, select = c("day", "dent.adult","dent.inf"),
                              host_para$rep == trails[i])
 }

@@

\begin{figure}[htbp]
\centering
\begin{subfigure}[b]{\linewidth}
<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=  
   mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
     cbind(mixed_nopara_data[[i]], group=paste("Trial", trails[i]))
   }))

 ggplot(mixed_nopara_data_combined, aes(x=day)) + 
   geom_line(aes(y=dent.adult, group=interaction(group, "dent.adult"))) + 
   geom_line(aes(y=lum.adult, color="blue", group=interaction(group, "lum.adult"))) +
   scale_x_continuous(limits = c(0, 52)) +
   scale_y_continuous(limits = c(0, 175)) +
   facet_grid(. ~ group, scales = "free", space = "free") +
   theme(
     strip.text = element_text(size=10),
     axis.text.x=element_text(color="black"),
     axis.ticks.x=element_line(color="black"),
     axis.text.y=element_text(color="black"),
     axis.ticks.y=element_line(color="black"),
     axis.title.x=element_blank(),
     axis.title.y=element_blank(),
     legend.position="none",
     panel.spacing = unit(0, "lines"), # Adjust spacing between panels
     strip.background = element_blank(), # Optional: Remove facet label background
     strip.placement = "outside"
   ) +
   labs(title="Daphnia density in Non-parasitized Dynamics") +
   guides(colour=guide_legend(title="Species"))
@
\caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in the experiment}
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=  
   mixed_para_data_combined <- do.call(rbind, lapply(1:length(mixed_para_data), function(i) {
     cbind(mixed_para_data[[i]], group=paste("Trial", trails[i]))
   }))

 ggplot(mixed_para_data_combined, aes(x=day)) + 
   geom_line(aes(y=dentadult, group=interaction(group, "dentadult"))) + 
   geom_line(aes(y=lumadult, color="blue", group=interaction(group, "lumadult"))) +
    geom_line(aes(y=luminf,color="green", group=interaction(group, "luminf"))) + 
   geom_line(aes(y=dentinf, color="red", group=interaction(group, "dentinf"))) +

   scale_x_continuous(limits = c(0, 52)) +
   scale_y_continuous(limits = c(0, 175)) +
   facet_grid(. ~ group, scales = "free", space = "free") +
   theme(
     strip.text = element_text(size=10),
     axis.text.x=element_text(color="black"),
     axis.ticks.x=element_line(color="black"),
     axis.text.y=element_text(color="black"),
     axis.ticks.y=element_line(color="black"),
     axis.title.x=element_blank(),
     axis.title.y=element_blank(),
     legend.position="none",
     panel.spacing = unit(0, "lines"), # Adjust spacing between panels
     strip.background = element_blank(), # Optional: Remove facet label background
     strip.placement = "outside"
   ) +
   labs(title="Daphnia density in Parasitized Dynamics") +
   guides(colour=guide_legend(title="Species"))
@
\caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in the experiment}
\end{subfigure}%
\end{figure}

@@

{
\small
\begin{align}
&d S(t) = r F(t)\left(S(t)+\xi I(t)\right)dt - \theta_{S}S(t) dt - p f_{S} S(t) Pdt - \delta S(t)dt + S(t) d\zeta_{S},\\
&d I(t) = p f_{S} S(t) Pdt - \theta_{I} I(t)dt - \delta I(t)dt + I(t) d\zeta_{I},\\
&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma F(t) \left(S(t)+\xi I(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F,\\
&d P(t) = \beta d I(t)dt - f_{S} S(t) P(t)dt - f_{I} I(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
&d\zeta_{S} \sim N\left(0, \sigma_{S}^{2}dt\right),\\
&d\zeta_{I} \sim N\left(0, \sigma_{I}^{2}dt\right),\\
&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}dt\right),\\
&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}dt\right).
\end{align}
}

@@

{
\small
\begin{align}
% \begin{split}
&d S_n(t) = r_n F(t) \left(S_n(t)+\xi I_n(t)\right)dt - \theta_{S_n}S_n(t) dt - p_n f_{S_n} S_n(t) P(t)dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n},\\
% \end{split}
% \\
&d I_n(t) = p_n f_{S_n} S_n(t) Pdt - \theta_{I_n} I_n(t)dt - \delta I_n(t)dt + I_n d\zeta_{I_n},\\
% \begin{split}
&d S_i(t) = r_i F(t)\left(S_i(t)+\xi I_i(t)\right)dt - \theta_{S_i} S_i(t) dt - p_i f_{S_i} S_i(t) P(t)dt -  \delta S_i(t)dt + S_i(t) d\zeta_{S_i},\\
% \end{split}
% \\
&d I_i(t) = p_i f_{S_i} S_i(t) P(t)dt - \theta_{I_i} I_i(t)dt - \delta I_i(t)dt + I_i(t) d\zeta_{I_i}(t),\\
\begin{split}
&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma_n F(t) \left(S_n(t)+\xi I_n(t)\right) dt - \gamma_i F(t) \left(S_i(t)+\xi I_i(t)\right) dt\\
&\quad \quad \quad- \delta F(t) dt + \mu dt + F(t) d\zeta_F,\\
\end{split}
\\
\begin{split}
&d P(t) = \beta_n d_n I_n(t)dt + \beta_i d_i I_i(t)dt - f_{S_n} S_n(t) P(t)dt - f_{I_n} I_n(t) P(t)dt - f_{S_i} S_i(t) P(t)dt \\
&\quad \quad \quad- f_{\mathrm{I}_i} I_i(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
\end{split}
\\
&d\zeta_{S_n} \sim N\left(0, \sigma_{S_n}^{2}dt\right),\\
&d\zeta_{I_n} \sim N\left(0, \sigma_{I_n}^{2}dt\right),\\
&d\zeta_{S_i} \sim N\left(0, \sigma_{S_i}^{2}dt\right),\\
&d\zeta_{I_i} \sim N\left(0, \sigma_{I_i}^{2}dt\right),\\
&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}dt\right),\\
&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}dt\right).
\end{align}
}

@@

$\gamma_n$ & Native host consumption rate&$10^{-4} \cdot L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gamman,2)}\\
$\gamma_i$ & Invsaive host consumption rate&$10^{-4} \cdot L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gammai,2)}\\

@@

$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSn,2)}\\
$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$10^{-4} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSi,2)}\\
$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIn,2)}\\
$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIi,2)}\\
$\sigma_{F}$ & Standard deviation of brownian motion of alga&$10^{-1} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigF,2)}\\
$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$10^{-2} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigP,2)}\\

@@

\begin{figure}
\begin{center}
%%%%% SEAIR diagram
\resizebox{8cm}{!}{
\begin{tikzpicture}[
  square/.style={rectangle, draw=black, minimum width=1cm, minimum height=1cm, rounded corners=.1cm, fill=blue!8},
  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=1cm,inner sep=0cm},
  rhombus/.style={diamond, draw=black, minimum width=0.1cm, minimum height=0.1cm, fill=purple!8,aspect = 1},
  bendy/.style={bend left=10},
  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
]
\tikzset{>={}}; % this is needed to implement the arrow redefinition

@@

\begin{align}
&d S_n(t)=r_n S_n(t) dt- \theta_{S_n}S_n(t) dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n}, \\
&d S_i(t)=r_i S_i(t) dt- \theta_{S_i} S_i(t) dt -\delta S_i(t)dt + S_i(t) d\zeta_{S_i}, \\
&dF(t)=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n \cdot dt - \gamma_i F S_i \cdot dt -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F.\\
&\zeta_F \sim N \left( 0, \sigma_F^2dt\right) \\
&\zeta_{S_n} \sim N \left( 0, \sigma_{S_n}^2dt  \right) \\
&\zeta_{S_i} \sim N \left( 0, \sigma_{S_i}^2dt \right) 
\end{align}

As a adaption form of SIRPF model, the upper equations show SF2 model, with the meaning of each parameter is the same as the SIRPF model. The only difference is that the SF and SF2 model do not have the parasitic term. While if we only keep one species in SF2 model, the model will become SF model. Utilizing the proposed model facilitates a rigorous way into the influence of algal density on the population dynamics of Daphnia, particularly under conditions of potential food scarcity. This analytical approach allows for the testing of a specific hypothesis: the critical role of algal density in determining the population trajectory of Daphnia. Should the model demonstrate a substantial fit to the observed data, it would substantiate the hypothesis that algal density is a pivotal factor in the observed decline of the Daphnia population, thereby suggesting food limitation as a primary driver. Conversely, a poor fit of the model to the data may indicate that factors other than algal density predominantly govern the population dynamics of Daphnia. In such a scenario, it becomes imperative to explore alternative variables that could significantly influence the population density of Daphnia.\\

@@

 ratio_SF_i = SF_invasive_estimate[length(SF_invasive_estimate)]/SF_invasive_estimate[4]
 ratio_SF_n = ratio_SF_i
 ratio_SF2_i = SF2_estimate[11]/SF2_estimate[8]
 ratio_SF2_n = SF2_estimate[10]/SF2_estimate[7]
 ratio_SIRPF_is = SIRPF_invasive_estimate[13]/SIRPF_invasive_estimate[8]
 ratio_SIRPF_ii = SIRPF_invasive_estimate[13] * SIRPF_invasive_estimate[14]/SIRPF_invasive_estimate[9]
 ratio_SIRPF_ns = SIRPF_native_estimate[13]/SIRPF_native_estimate[8]
 ratio_SIRPF_ni = SIRPF_native_estimate[13] * SIRPF_native_estimate[14]/SIRPF_native_estimate[9]
 ratio_SIRPF2_ns = SIRPF2_estimate[20]/SIRPF2_estimate[10]
 ratio_SIRPF2_ni = SIRPF2_estimate[20] *  SIRPF2_estimate[2]/SIRPF2_estimate[12]
 ratio_SIRPF2_is = SIRPF2_estimate[21]/SIRPF2_estimate[11]
 ratio_SIRPF2_ii = SIRPF2_estimate[21] *  SIRPF2_estimate[2]/SIRPF2_estimate[14]
@
When constructing statistical models to analyze ecological data, the specification of parameters plays a crucial role in shaping the model's complexity and interpretability. In a scenario with 10 units of observation, the decision to assign a parameter as shared across units or unique to each one has significant implications for the model's dimensionality. For instance, designating a parameter as shared across all units incrementally increases the model dimension by one, whereas categorizing a parameter as unique to each unit multiplies the dimensionality by the number of units, which in this case results in an increase by 10 for single species Non-parasitized dynamics, 9 for mixed species Non-parasitized dynamics and Parasitized dynamics with the invasive species of Daphnia, and 8 for the rest dynamics.\\

@@

By leveraging the estimated parameters from the most suitable model, we can conduct a nuanced numerical analysis of the impact of food resources and parasitism on the interactions between local and invasive host species. The survivability ratio($ \lambda = \frac{r}{\theta}$) of growth to mortality rates is a useful metric for assessing the health of a population, as it provides a measure of the population's ability to sustain itself in the face of environmental pressures \citep{ma2021unified}. This approach, grounded in ecological principles, enables a detailed examination of the survival strategies of the hosts in response to external challenge. The survivability ratio of native and invasive Daphnia in single species Non-parasitized Dynamics are \Sexpr{mysignif(ratio_SF_n,2)} and \Sexpr{mysignif(ratio_SF_i,2)} respectivly. While if we mix two species of Daphnia into a Non-parasitized dynamics, the ratio drops down to \Sexpr{mysignif(ratio_SF2_n,2)} for native Daphnia and to \Sexpr{mysignif(ratio_SF2_i,2)} for invasive species. In such scenario, we observe significant fluctuations in the population density of Daphnia. These oscillations are likely attributable to the constraints imposed by limited food resources within the environment. Such environmental limitations typically lead to variations in population dynamics. The observed variability in the survivability ratio can be understood within this assumption. At the same time, it is important to note that the relative survivability between the two Daphnia species studied remains consistent, despite an overall simultaneous decrease in their population densities. This suggests that while environmental factors like food scarcity impact the survivability of both species, they do so in a proportionally similar manner, maintaining the relative size of survivability between them. And similar results can be found in the Parasitized dynamics of invasive species of Daphnia, where the survivability ratio of susceptible native and invasive Daphnia are \Sexpr{mysignif(ratio_SIRPF_ns,2)} and \Sexpr{mysignif(ratio_SIRPF_is,2)} for single species dynamics, and \Sexpr{mysignif(ratio_SIRPF2_ns,2)}, \Sexpr{mysignif(ratio_SIRPF2_is,2)} for native and invasive population in mixed species dynamics respectively.\\

>>>>

<<Setup, include=FALSE,echo=FALSE,results=hide>>=
 library(plyr)
 library(reshape2)
library(patchwork)
 library(magrittr)
 library(foreach)
 library(readxl)
 library(doParallel)
 registerDoParallel(cores=36)
 library(pomp)
 library(panelPomp)
 library(ggplot2)

@@

 #---------------------------Host Daphnia----------------------------------------------------------------------
 #----------------load data---------------------------------------------------------------
load("data/raw_data.RData")
 #-----------------------------Host with para----------------------------------------------------------------------------
 host_para <- Mesocosm_data[101:180, ]
 host_para <- subset(host_para, select = c(rep, day, dent.adult,dent.inf ))
 host_para <- host_para[80: 1, ]
 #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
 host_para$day = (host_para$day - 1) * 5 + 7
 host_para_data = list()
 trails = c("K","L","M","N","O","P","Q","S")
 for (i in 1: length(trails)){
   host_para_data[[i]]=subset(host_para, select = c("day", "dent.adult","dent.inf"),
                              host_para$rep == trails[i])
 }

@@

\begin{figure}[H]
\centering
<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=16, height=10>>=  
   mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
     cbind(mixed_nopara_data[[i]], group=paste("Trial", trails[i]))
   }))
    mixed_para_data_combined <- do.call(rbind, lapply(1:length(mixed_para_data), function(i) {
     cbind(mixed_para_data[[i]], group=paste("Trial", trails[i]))
   }))

p1<- ggplot(mixed_nopara_data_combined, aes(x=day)) + 
   geom_line(aes(y=dent.adult, group=interaction(group, "dent.adult"))) + 
   geom_line(aes(y=lum.adult, color="blue", group=interaction(group, "lum.adult"))) +
   scale_x_continuous(limits = c(0, 52)) +
   scale_y_continuous(limits = c(0, 175)) +
   facet_grid(. ~ group, scales = "fixed", space = "free") +
   theme(
     strip.text = element_text(size=10),
     axis.text.x=element_text(color="black"),
     axis.ticks.x=element_line(color="black"),
     axis.text.y=element_text(color="black"),
     axis.ticks.y=element_line(color="black"),
     axis.title.x=element_blank(),
     axis.title.y=element_blank(),
     legend.position="none",
     panel.spacing = unit(0, "lines"), # Adjust spacing between panels
     strip.background = element_blank(), # Optional: Remove facet label background
     strip.placement = "outside"
   ) +
   labs(title="Daphnia density in Non-parasitized Dynamics") +
   guides(colour=guide_legend(title="Species"))

p2<- ggplot(mixed_para_data_combined, aes(x=day)) + 
   geom_line(aes(y=dentadult, group=interaction(group, "dentadult"))) + 
   geom_line(aes(y=lumadult, color="blue", group=interaction(group, "lumadult"))) +
    geom_line(aes(y=luminf,color="green", group=interaction(group, "luminf"))) + 
   geom_line(aes(y=dentinf, color="red", group=interaction(group, "dentinf"))) +

   scale_x_continuous(limits = c(0, 52)) +
   scale_y_continuous(limits = c(0, 175)) +
   facet_grid(. ~ group, scales = "fixed", space = "free") +
   theme(
     strip.text = element_text(size=10),
     axis.text.x=element_text(color="black"),
     axis.ticks.x=element_line(color="black"),
     axis.text.y=element_text(color="black"),
     axis.ticks.y=element_line(color="black"),
     axis.title.x=element_blank(),
     axis.title.y=element_blank(),
     legend.position="none",
     panel.spacing = unit(0, "lines"), # Adjust spacing between panels
     strip.background = element_blank(), # Optional: Remove facet label background
     strip.placement = "outside"
   ) +
   labs(title="Daphnia density in Parasitized Dynamics") +
   guides(colour=guide_legend(title="Species"))
combined_plot <- p1 / p2 
combined_plot
@
\caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in different experiment. The black lines represent the Daphnia dentifera density, the red lines show the Daphnia lumholtzi density, the green lines are the infected Daphnia lumholtzi density, and the blue lines indicate the infected Daphnia dentifera density.}
\end{figure}

@@

{
\small
\begin{align}
&d S(t) = r F(t)\left(S(t)+\xi I(t)\right)dt - \theta_{S}S(t) dt - p f_{S} S(t) Pdt - \delta S(t)dt + S(t) d\zeta_{S},\\
&d I(t) = p f_{S} S(t) Pdt - \theta_{I} I(t)dt - \delta I(t)dt + I(t) d\zeta_{I},\\
&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma F(t) \left(S(t)+\xi I(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F,\\
&d P(t) = \beta \theta_I I(t)dt - f_{S} S(t) P(t)dt - f_{I} I(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
&d\zeta_{S} \sim N\left(0, \sigma_{S}^{2}dt\right),\\
&d\zeta_{I} \sim N\left(0, \sigma_{I}^{2}dt\right),\\
&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}dt\right),\\
&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}dt\right).
\end{align}
}

@@

{
\small
\begin{align}
% \begin{split}
&d S_n(t) = r_n F(t) \left(S_n(t)+\xi I_n(t)\right)dt - \theta_{S_n}S_n(t) dt - p_n f_{S_n} S_n(t) P(t)dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n},\\
% \end{split}
% \\
&d I_n(t) = p_n f_{S_n} S_n(t) Pdt - \theta_{I_n} I_n(t)dt - \delta I_n(t)dt + I_n d\zeta_{I_n},\\
% \begin{split}
&d S_i(t) = r_i F(t)\left(S_i(t)+\xi I_i(t)\right)dt - \theta_{S_i} S_i(t) dt - p_i f_{S_i} S_i(t) P(t)dt -  \delta S_i(t)dt + S_i(t) d\zeta_{S_i},\\
% \end{split}
% \\
&d I_i(t) = p_i f_{S_i} S_i(t) P(t)dt - \theta_{I_i} I_i(t)dt - \delta I_i(t)dt + I_i(t) d\zeta_{I_i}(t),\\
\begin{split}
&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma_n F(t) \left(S_n(t)+\xi I_n(t)\right) dt - \gamma_i F(t) \left(S_i(t)+\xi I_i(t)\right) dt\\
&\quad \quad \quad- \delta F(t) dt + \mu dt + F(t) d\zeta_F,\\
\end{split}
\\
\begin{split}
&d P(t) = \beta_n \theta_{I_n} I_n(t)dt + \beta_i \theta_{I_i} I_i(t)dt - f_{S_n} S_n(t) P(t)dt - f_{I_n} I_n(t) P(t)dt - f_{S_i} S_i(t) P(t)dt \\
&\quad \quad \quad- f_{\mathrm{I}_i} I_i(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
\end{split}
\\
&d\zeta_{S_n} \sim N\left(0, \sigma_{S_n}^{2}dt\right),\\
&d\zeta_{I_n} \sim N\left(0, \sigma_{I_n}^{2}dt\right),\\
&d\zeta_{S_i} \sim N\left(0, \sigma_{S_i}^{2}dt\right),\\
&d\zeta_{I_i} \sim N\left(0, \sigma_{I_i}^{2}dt\right),\\
&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}dt\right),\\
&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}dt\right).
\end{align}
}

@@

$\gamma_n$ & Native host consumption rate&$10^{-4} \cdot L \cdot individual^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gamman,2)}\\
$\gamma_i$ & Invsaive host consumption rate&$10^{-4} \cdot L \cdot individual^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gammai,2)}\\

@@

$\sigma_{S_n}$ & Standard deviation of Brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSn,2)}\\
$\sigma_{S_i}$ & Standard deviation of Brownian motion of susceptible invasive &$10^{-4} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSi,2)}\\
$\sigma_{I_n}$ & Standard deviation of Brownian motion of infected native&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIn,2)}\\
$\sigma_{I_i}$ & Standard deviation of Brownian motion of infected invasive&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIi,2)}\\
$\sigma_{F}$ & Standard deviation of Brownian motion of alga&$10^{-1} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigF,2)}\\
$\sigma_{P}$ &Standard deviation of Brownian motion of parasite &$10^{-2} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigP,2)}\\

@@

\begin{figure}[H]
\begin{center}
%%%%% SEAIR diagram
\resizebox{8cm}{!}{
\begin{tikzpicture}[
  square/.style={rectangle, draw=black, minimum width=1cm, minimum height=1cm, rounded corners=.1cm, fill=blue!8},
  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=1cm,inner sep=0cm},
  rhombus/.style={diamond, draw=black, minimum width=0.1cm, minimum height=0.1cm, fill=purple!8,aspect = 1},
  bendy/.style={bend left=10},
  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
]
\tikzset{>={}}; % this is needed to implement the arrow redefinition

@@

\begin{align}
&d S_n(t)=r_n S_n(t) dt- \theta_{S_n}S_n(t) dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n}, \\
&d S_i(t)=r_i S_i(t) dt- \theta_{S_i} S_i(t) dt -\delta S_i(t)dt + S_i(t) d\zeta_{S_i}, \\
&dF(t)=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n(t) \cdot dt - \gamma_i F(t) S_i(t) \cdot dt -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F.\\
&\zeta_F \sim N \left( 0, \sigma_F^2dt\right) \\
&\zeta_{S_n} \sim N \left( 0, \sigma_{S_n}^2dt  \right) \\
&\zeta_{S_i} \sim N \left( 0, \sigma_{S_i}^2dt \right) 
\end{align}

As a adaption form of SIRPF model, the upper equations show SF2 model, with the meaning of each parameter is the same as the SIRPF model. The only difference is that the SF and SF2 model do not have the parasitic term. While if we only keep one species in SF2 model, the model will become SF model. Utilizing the proposed model facilitates a rigorous way into the influence of algal density on the population dynamics of Daphnia, particularly under conditions of potential food scarcity. This analytical approach allows for the testing of a specific hypothesis: the critical role of algal density in determining the population trajectory of Daphnia. Should the model demonstrate a substantial fit to the observed data, it would substantiate the hypothesis that algal density is a pivotal factor in the observed decline of the Daphnia population, thereby suggesting food limitation as a primary driver \citep{xiaotong2018}. Conversely, a poor fit of the model to the data may indicate that factors other than algal density predominantly govern the population dynamics of Daphnia. In such a scenario, it becomes imperative to explore alternative variables that could significantly influence the population density of Daphnia.\\

@@

 ratio_SF_i = SF_invasive_estimate[length(SF_invasive_estimate)]/SF_invasive_estimate[4]
 ratio_SF_n = SF_native_estimate[length(SF_native_estimate)]/SF_native_estimate[4]
 ratio_SF2_i = SF2_estimate[11]/SF2_estimate[8]
 ratio_SF2_n = SF2_estimate[10]/SF2_estimate[7]
 ratio_SIRPF_is = SIRPF_invasive_estimate[13]/SIRPF_invasive_estimate[8]
 ratio_SIRPF_ii = SIRPF_invasive_estimate[13] * SIRPF_invasive_estimate[14]/SIRPF_invasive_estimate[9]
 ratio_SIRPF_ns = SIRPF_native_estimate[13]/SIRPF_native_estimate[8]
 ratio_SIRPF_ni = SIRPF_native_estimate[13] * SIRPF_native_estimate[14]/SIRPF_native_estimate[9]
 ratio_SIRPF2_ns = mif.estimate$rn/mif.estimate$theta_Sn
 ratio_SIRPF2_ni = SIRPF2_estimate[20] *  SIRPF2_estimate[2]/SIRPF2_estimate[12]
 ratio_SIRPF2_is = SIRPF2_estimate[21]/SIRPF2_estimate[11]
 ratio_SIRPF2_ii = SIRPF2_estimate[21] *  SIRPF2_estimate[2]/SIRPF2_estimate[14]

 save(SF_invasive_estimate=SF_invasive_estimate,SF_native_estimate=SF_native_estimate,SF2_estimate=SF2_estimate,SIRPF_invasive_estimate=SIRPF_invasive_estimate,SIRPF_native_estimate=SIRPF_native_estimate,SIRPF2_estimate=SIRPF2_estimate,file = "SIRPF_model_estimate.RData")
@
When constructing statistical models to analyze ecological data, the specification of parameters plays a crucial role in shaping the model's complexity and interpretability. In a scenario with 10 units of observation, the decision to assign a parameter as shared across units or unique to each one has significant implications for the model's dimensionality. For instance, designating a parameter as shared across all units incrementally increases the model dimension by one, whereas categorizing a parameter as unique to each unit multiplies the dimensionality by the number of units, which in this case results in an increase by 10 for single species Non-parasitized dynamics, 9 for mixed species Non-parasitized dynamics and Parasitized dynamics with the invasive species of Daphnia, and 8 for the rest dynamics.\\

@@

By leveraging the estimated parameters from the most suitable model, we can conduct a nuanced numerical analysis of the impact of food resources and parasitism on the interactions between local and invasive host species. The survivability ratio($ \lambda = \frac{r}{\theta}$) of growth to mortality rates is a useful metric for assessing the health of a population, as it provides a measure of the population's ability to sustain itself in the face of environmental pressures \citep{ma2021unified}. This approach, grounded in ecological principles, enables a detailed examination of the survival strategies of the hosts in response to external challenge. The survivability ratio of native and invasive Daphnia in single species Non-parasitized Dynamics are \Sexpr{mysignif(ratio_SF_n,3)} and \Sexpr{mysignif(ratio_SF_i,3)} respectivly. While if we mix two species of Daphnia into a Non-parasitized dynamics, the ratio drops down to \Sexpr{mysignif(ratio_SF2_n,3)} for native Daphnia and to \Sexpr{mysignif(ratio_SF2_i,3)} for invasive species. In such scenario, we observe significant fluctuations in the population density of Daphnia. These oscillations are likely attributable to the constraints imposed by limited food resources within the environment. Such environmental limitations typically lead to variations in population dynamics. The observed variability in the survivability ratio can be understood within this assumption. At the same time, it is important to note that the relative survivability between the two Daphnia species studied remains consistent, despite an overall simultaneous decrease in their population densities. This suggests that while environmental factors like food scarcity impact the survivability of both species, they do so in a proportionally similar manner, maintaining the relative size of survivability between them. And similar results can be found in the Parasitized dynamics of invasive species of Daphnia, where the survivability ratio of susceptible native and invasive Daphnia are \Sexpr{mysignif(ratio_SIRPF_ns,3)} and \Sexpr{mysignif(ratio_SIRPF_is,3)} for single species dynamics, and \Sexpr{mysignif(ratio_SIRPF2_ns,3)}, \Sexpr{mysignif(ratio_SIRPF2_is,3)} for native and invasive population in mixed species dynamics respectively.\\
